---
title: "Homework 7"
author: "Alicja Dutkiewicz, Irina Gurewitsch, João Esteves"
date: "04/11/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(igraph)
library(purrr)
library(data.table)
library(R.utils)
library(seqinr) # swap
library(knitr) #tables
library(kableExtra) # tables
```

## Exercise 7.1

#### (1)

```{r}
# Safer version of sample, fixes results for length 1
resample <- function(x, ...) x[sample.int(length(x), ...)]

genConfModel <- function(degs) {
  edges <- list()
  
  repeat {
    # Find unconnected stubs
    non_zero_inds <- which(degs > 0)
    if (length(non_zero_inds) < 2 & length(which(degs >= 2)) == 0) {
      break
    }
  
    # Randomly pick a stub
    stub1_ind <- resample(non_zero_inds, 1)
    degs[stub1_ind] <- degs[stub1_ind] - 1
  
    # Randomly pick a stub to connect the previous one to
    non_zero_inds <- which(degs > 0)
    stub2_ind <- resample(non_zero_inds, 1)
    degs[stub2_ind] <- degs[stub2_ind] - 1
    
    edges <- append(edges, list(list(stub1_ind, stub2_ind)))
  }
  
  links <- data.frame(t(sapply(edges, c)))
  
  return (links)
}
```

#### (2)

TODO

#### (3)

```{r}
degs1 <- c(2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2)
degs2 <- c(7, 4, 2, 1, 1, 1, 1, 1, 1, 1, 1, 1)
```

| k<sub>1</sub> self-loops | k<sub>2</sub> self-loops | k<sub>1</sub> repeated edges | k<sub>2</sub> repeated edges |
|---------------|---------------|-------------------|-------------------|
| 4             | 3             | 0                 | 3                 |
| 1             | 3             | 4                 | 3                 |
| 1             | 2             | 0                 | 0                 |
| 1             | 1             | 0                 | 3                 |
| 1             | 2             | 2                 | 3                 |
| 0             | 3             | 0                 | 2                 |
| 0             | 1             | 2                 | 3                 |
| 1             | 3             | 4                 | 2                 |
| 0             | 4             | 4                 | 4                 |
| 0             | 2             | 2                 | 2                 |
| 1             | 2             | 0                 | 4                 |
| 0             | 1             | 0                 | 2                 |
| 0             | 3             | 0                 | 2                 |
| 1             | 3             | 4                 | 3                 |
| 1             | 3             | 0                 | 2                 |
| 2             | 2             | 0                 | 2                 |
| 1             | 5             | 0                 | 3                 |
| 2             | 2             | 0                 | 2                 |
| 0             | 4             | 2                 | 3                 |
| 1             | 2             | 0                 | 2                 |
| 0             | 3             | 0                 | 2                 |
| 1             | 1             | 6                 | 2                 |
| 2             | 1             | 0                 | 2                 |
| 1             | 4             | 0                 | 3                 |
| 0             | 2             | 0                 | 4                 |
| 1             | 3             | 0                 | 4                 |
| 0             | 2             | 0                 | 4                 |
| 1             | 2             | 4                 | 4                 |
| 1             | 3             | 0                 | 2                 |
| 0             | 3             | 0                 | 5                 |

The average number of self-loops in k1 is 0.83 while for k2 it's 2.5. The average number of repeated edges in k1 is 1.13 while for k2 it's 2.73.

## Exercise 7.2

#### (4)

Below, the function *genRepConfModel()* generates a graph from a given degree sequence via the Repeated Configuration Model.

```{r eval=FALSE}
degs1 <- c(2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2)
degs2 <- c(7, 4, 2, 1, 1, 1, 1, 1, 1, 1, 1, 1)

# Safer version of sample, fixes results for length 1
resample <- function(x, ...) x[sample.int(length(x), ...)]

genRepConfModel <- function(degs) {
  edges <- list()
  
  repeat {
    # Find unconnected stubs
    non_zero_inds <- which(degs > 0)
    if (length(non_zero_inds) < 2 & length(which(degs >= 2)) == 0) {
      break
    }
  
    # Randomly pick a stub
    stub1_ind <- resample(non_zero_inds, 1)
    degs[stub1_ind] <- degs[stub1_ind] - 1
  
    # Randomly pick a stub to connect the previous one to
    non_zero_inds <- which(degs > 0)
    stub2_ind <- resample(non_zero_inds, 1)
    degs[stub2_ind] <- degs[stub2_ind] - 1
    
    # These conditional stops are the only differences from exercise 1.1
    if (stub1_ind == stub2_ind) {
      stop('Self-loop detected')
    }
    if (list(list(stub1_ind, stub2_ind)) %in% edges | list(list(stub2_ind, stub1_ind)) %in% edges) {
      stop('Repeated edge detected')
    }
    
    edges <- append(edges, list(list(stub1_ind, stub2_ind)))
  }
  
  links <- data.frame(t(sapply(edges, c)))
  
  return (links)
}

drawNewRepCM <- function(degs) {
  success <- FALSE
  i <- 0
  while (!success) {
    i <- i + 1
    printf('Generating graph: attempt %d\n', i)
    links <- tryCatch({
      genRepConfModel(degs)
    }, error = function(msg) {
      message('Restarting program. Reason:')
      message(msg)
      message('\n')
      return (NA)
    })
    if (!is.na(links)) {
      success <- TRUE
    }
  }
  network <- graph_from_data_frame(d=links, directed=FALSE)
  plot(network)
  return (i)
}

drawNewRepCM(degs1)
```

The code below generates 30 graphs via the function above with the goal of measuring the necessary amount of restarts each time.

```{r eval=FALSE}
restarts <- as.integer(list())
for (i in 1:30) {
  # restarts = attempts - 1
  restarts <- append(restarts, drawNewRepCM(degs1) - 1)
}
printf('Restarts: ')
for (r in restarts) {
  printf('%d ', r)
}
printf('Average restarts: %f\n', mean(restarts))
```

The results were that the average number of restarts required for k<sub>1</sub> were 2.133333 and for k<sub>2</sub> were 669.300. In addition, it was observed that the majority of these restarts were due to self-loops.
Generalizing this example, it means that degree distributions including elements with higher degrees tend to fail more. My explanation for this is that, as the vertices with smaller degrees get picked and crossed out the list, fewer vertices (with high degrees) remain and thus it's more likely they'll connect with themselves. In addition, having vertices with high degrees means they'll take longer to be picked for all the edges they can have, so they have more chances to repeat connections.

## Exercise 7.3

#### (5)

In the code chunk below you'll find functions to obtain, respectively, the degree of a node, the local clustering coefficient of a node and the average clustering coefficient. The time complexity for the former is O(|V|<sup>3</sup>).

Note that for this exercise the graph representation is now an adjacency matrix rather than a list of edges.

```{r}
# Time: O(|V|)
getNodeDegree <- function(adjMat, node) {
  sum(adjMat[node,])
}

# Time: O(|V| + |V|^2) = O(|V|^2)
getLocalClusterCoef <- function(adjMat, node) {
  nodeDeg <- getNodeDegree(adjMat, node)
  if (nodeDeg <= 1) {
    return (0)
  }
  triangles <- 0
  for (n1 in setdiff(1:nrow(adjMat), node)) {
    for (n2 in setdiff(1:nrow(adjMat), c(node, n1))) {
      triangles <- triangles + adjMat[node, n1] * adjMat[n1, n2] * adjMat[n2, node]
    }
  }
  return (triangles / (nodeDeg * (nodeDeg - 1)))
}

# Time: O(|V|*(|V|^2)) = O(|V|^3)
getAvgClusterCoef <- function(adjMat) {
  sum <- 0
  for (node in 1:nrow(adjMat)) {
    sum <- sum + getLocalClusterCoef(adjMat, node)
  }
  return (sum / nrow(adjMat))
}
```

#### (6)

The code below defines functions for the elementary rewiring step and for the local rewiring algorithm.

```{r}
rewireStep <- function(adjMat) {
  edges <- which(adjMat != 0, arr.ind=TRUE)
  repeat {
    edge1 <- edges[sample(nrow(edges), size=1), ]
    edge2 <- edges[sample(nrow(edges), size=1), ]
    # accept edges if they don't share any nodes and if the new edges don't already exist
    if (length(intersect(edge1, edge2)) == 0
        & adjMat[edge1[1], edge2[2]] == 0 & adjMat[edge2[1], edge1[2]] == 0) {
      break
    }
  }
  
  # Remove current edges
  adjMat[edge1[1], edge1[2]] <- adjMat[edge1[2], edge1[1]] <- 0
  adjMat[edge2[1], edge2[2]] <- adjMat[edge2[2], edge2[1]] <- 0
  
  # Insert new swapped edges
  adjMat[edge1[1], edge2[2]] <- adjMat[edge2[2], edge1[1]] <- 1
  adjMat[edge2[1], edge1[2]] <- adjMat[edge1[2], edge2[1]] <- 1
  
  return (adjMat)
}

# its: Iterations
localRewiring <- function(adjMat, its) {
  for (i in 1:its) {
    adjMat <- rewireStep(adjMat)
  }
  return (adjMat)
}
```

The code below runs the local rewiring algorithm and measures the obtained graphs' average cluster coefficients to generate a table.

```{r}
# Graph with 4 disconnected clusters
g1 <- matrix(0, nrow=12, ncol=12)
g1Edges <- data.frame('src' = 1:12, 'target' = c(2,3,1, 5,6,4, 8,9,7, 11,12,10))
g1[cbind(g1Edges$src, g1Edges$target)] <- g1[cbind(g1Edges$target, g1Edges$src)] <- 1

# Circular graph
g2 <- matrix(0, nrow=12, ncol=12)
g2Edges <- data.frame('src' = 1:12, 'target' = (1:12 %% 12) + 1)
g2[cbind(g2Edges$src, g2Edges$target)] <- g2[cbind(g2Edges$target, g2Edges$src)] <- 1

g1_avg <- getAvgClusterCoef(g1)
g2_avg <- getAvgClusterCoef(g2)

g1r5 <- g1r40 <- g2r5 <- g2r40 <- c()
for (i in 1:10) {
  g1_rewired5 <- localRewiring(g1, its=5)
  g2_rewired5 <- localRewiring(g2, its=5)
  
  g1_rewired5_avg <- getAvgClusterCoef(g1_rewired5)
  g2_rewired5_avg <- getAvgClusterCoef(g2_rewired5)
  
  g1_rewired40 <- localRewiring(g1, its=40)
  g2_rewired40 <- localRewiring(g2, its=40)
  
  g1_rewired40_avg <- getAvgClusterCoef(g1_rewired40)
  g2_rewired40_avg <- getAvgClusterCoef(g2_rewired40)
  
  g1r5 <- append(g1r5, g1_rewired5_avg)
  g1r40 <- append(g1r40, g1_rewired40_avg)
  g2r5 <- append(g2r5, g2_rewired5_avg)
  g2r40 <- append(g2r40, g2_rewired40_avg)
}

avgTable <- data.frame(Experiment = 1:10,
                       G1.5 = g1r5,
                       G2.5 = g2r5,
                       G1.40 = g1r40,
                       G2.40 = g2r40)
kable(avgTable,
      col.names = c('Experiment','G1','G2','G1','G2'),
      align=c('l','c','c','c','c')
      ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered"), full_width = F, position = 'left') %>%
  add_header_above(c('', '5 iterations' = 2, '40 iterations' = 2))
```

In addition to the table, the average cluster coefficients before rewiring for G1 and G2 are respectively 1.0 and 0.0.
As seen in the table, there aren't many differences between starting the algorithm on G1 or G2. This is because G1 can turn into G2 and vice-versa after just a few rewiring steps, so the results are quite random.